GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches

### 2025-05-15 response2reviewer.tex

\begin{center}
    \maketitle
\end{center}

%%%%%%%%%%%%%%%%
% cover letter %
%%%%%%%%%%%%%%%%
\thedate

Dear Editor:

Thank you for giving us the opportunity to submit a revised draft of the manuscript ``\thetitle'' for publication in the Journal of \thejournal. We appreciate the time and effort that editors and the reviewers dedicated to providing feedback on our manuscript and are grateful for the insightful comments and valuable improvements to our paper. We have incorporated the suggestions made by the reviewers. Those changes are highlighted within the manuscript. Please see below for a point-by-point response to the reviewers' comments and concerns. All page numbers refer to the revised manuscript file with tracked changes.

Thank you for your consideration. I am looking forward to hearing from you soon.

@@

    When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.
\end{reviewercomment}

\response{
    We sincerely appreciate the positive feedback from the editor and reviewers regarding our manuscript. 
    We are grateful for the opportunity to revise our work based on the valuable comments received.

    In this revision, we have:

    1. Carefully addressed all reviewer comments point-by-point;\\
    2. Especially, ...
    % 2. Provided detailed responses explaining each modification;\\
    % 3. Made all necessary revisions to improve manuscript clarity and technical accuracy;\\
    % 4. Included tracked changes in the revised manuscript for transparent review.

    These revisions have significantly strengthened the paper, and we appreciate the editors' and reviewers' time and constructive suggestions. 
    Please find our point-by-point responses to each reviewer's comment in the following sections.
}

@@

\response{
    We sincerely appreciate the reviewer's valuable feedback and constructive suggestions for improving our paper. 
    Regarding the request for additional experiments on instance segmentation and completion,
    We have enhanced the manuscript to better explain these design choices in the Methods, Results and  Discussion sections.
    Please find our detailed responses to each specific comment below.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Regarding the instance segmentation, I understand that this is not the main contribution of the paper. Still, it would be nice to test different approaches here to evaluate the robustness of the completion network to varying levels of wrong predictions. Along a similar line, I would have liked to see the completion network working on ground truth segmentation labels as this will disentangle errors coming from the instance segmentation net and the completion net.
\end{reviewercomment}

\response{
    We sincerely appreciate the reviewer's insightful comments regarding the evaluation of instance segmentation and its impact on the completion network. 
    We agree that analyzing the robustness of the completion network to segmentation errors is an interesting direction, and we already have conducted a similar preliminary experiments during the early stage of our studies.

    In our preliminary tests on a small subset of data, we visually compared the outputs of the completion network on using ground truth segmentation versus SoftGroup segmentation outputs. 
    The differences were negligible to the human eye and thus not mentioned in the manuscript.
    This is because our high-quality grape point clouds (captured in a controlled indoor environment with high-overlap scanning) exhibit clear boundaries and minimal noise. 
    Also, since our segmentation network (trained on manually annotated data using SoftGroup) achieves near-perfect performance (mAP@[.5:.95] of 0.995, representing AP50 of 0.999), making its outputs virtually indistinguishable from ground truth for our application.
    Moreover, we note that manual annotation of 3D data on 2D displays inevitably contains some human error, meaning even ``ground truth'' segmentations are not perfectly accurate.

    Given these observations, our current design - where each specialized network focuses on its respective task - represents the most efficient approach. 
    The segmentation network reliably handles instance separation, while the completion network specializes in partial point cloud completion.
    While artificially generating poor segmentation outputs to test completion robustness could be interesting, 
    It would require additional design to generate extra training data to cover various error patterns that might negatively impact network training.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Regarding the completion net, I would have liked to see ablation studies here as I believe this is the main  contribution of this work. For example, how does the number of cuts  influence the results of the completion metrics? How does the number of  output points in the centroid-based contour influence the snowflake net?
\end{reviewercomment}

\response{
    We thank the reviewer for the insightful suggestions regarding ablation studies for our completion network. 
    We would like to clarify that GrapeCPNet is not an entirely new network architecture, but rather a combination of features from two existing models (PoinTr and SnowFlakeNet). 
    This approach differs from traditional network design methodologies. 
    Therefore, performing the suggested ablation studies may not be practical or aligned with our main research objectives.

    However, when treating each network component as an independent module, our results already include comparative experiments between individual networks and the combined architecture. 
    As shown in Table 3, we compared the performance of: (1) PoinTr alone; (2) SnowFlakeNet alone; (3) Our combined GrapeCPNet. 
    The results demonstrate that the combined architecture outperforms either individual network.

    To support future research in this area, we also made our source code and training data of completion net publicly available. 
    This will allow other researchers to examine how preprocessing parameters affect the completion network's performance.

    We have also added the above discussion to the results section.
}

@@

\manuscript{
    this paper further validated the efficacy of the 3D phenotypic traits of grape by the proposed method.

    \added[id=R1]{
        Since GrapeCPNet combines features from two existing models (PoinTr and SnowFlakeNet) rather than introducing a completely new architecture, ablation studies may not align with our research focus.
        However, we conducted comparative experiments by treating each network as an independent module.
        As shown in Table~\ref{tbl:4}, we evaluated three approaches: (1) PoinTr alone, (2) SnowFlakeNet alone, and (3) our combined GrapeCPNet.
        The results show that GrapeCPNet achieves better performance than either individual network.
        The source code and training data are publicly available at \url{https://github.com/I3-Laboratory/GrapeCPNet}.
        This will enable interested researchers to examine how preprocessing parameters influence the network's performance.
    }

@@

    \textbf{3.3 Phenotypic Traits Calculation Results}
}

\manuscript{
    \textbf{Data availability}

    The \added[id=R1]{demo dataset, source codes} and model weight \added[id=R1]{of GrapeCPNet (completion network)} for the \replaced[id=R1]{academic}{test} usage is available \replaced[id=R1]{at}{here:} \url{https://github.com/I3-Laboratory/GrapeCPNet}. 
}

### 2025-05-15 main.tex
wei's updates

\section*{Data availability}
The \added[id=R1]{demo dataset, source codes} and model weight for the \replaced[id=R1]{academic}{test} usage is available \replaced[id=R1]{at}{here:} \url{https://github.com/I3-Laboratory/GrapeCPNet}. 
The whole dataset \added[id=R1]{and codes} used in this paper is available upon request.

>>>>

\section*{Data availability}
The \added[id=R1]{demo dataset, source codes} and model weight \added[id=R1]{of GrapeCPNet (completion network)} for the \replaced[id=R1]{academic}{test} usage is available \replaced[id=R1]{at}{here:} \url{https://github.com/I3-Laboratory/GrapeCPNet}.
